---
course: 'DS2003: Communicating with Data'
class title: 'Introduction to R: Part 1'
class number: '2'
output:
  html_document: default
  pdf_document: default
---
##############################
# R as a simple calculator
##############################

### Arithmetic operators
* Addition: +
* Subtraction: -
* Multiplication: *
* Division: /
* Exponentiation: ^ (e.g. 4^4 = 16)
* Modulo: %% (returns the remainder of the division of the number to the left by the number on its right; 
e.g. 5 %% 3 = 2)
* Square root: sqrt() (this a function vs an operator; e.g. sqrt(16) = 4)


```{r}
# Example: Addition

2 + 1

```

```{r}
# (s1) Addition - write an addition equation

x = 4+3

```

```{r}
# (s2) Subtraction - write a subtraction equation

y = 5 - 3

```

```{r}
# (s3) Multiplication - write a multiplication equation

z = 4 * 3

```

```{r}
# (s4) Division - write a division equation

t = 2/3

```

```{r}
# (s5) Exponentiation - write an exponentiation equation
g = 2^3

```

```{r}
# (s6) Modulo - write a modulo equation
o = 10 %% 10

```

### Remember PEMDAS (Parentheses, Exponents, Multiplication, Division, Addition, Subtraction)

```{r}
# Equation with PEMDAS

100/(27-2)

#versus

100/27-2
```

```{r}
# (s7) Write an equation where PEMDAS is applicable

y1 = g*(o-t)

#vs

y2 = (g*o)-t

```

#################################
# Logical (comparison) operators
#################################

* < for less than
* > for greater than
* <= for less than or equal to
* >= for greater than or equal to
* == for equal to each other
* != not equal to each other

```{r}
# Example: R will tell us if the expression is TRUE or FALSE
6 < 5


```

```{r}
# s(7b) Write an expression with a logical operator that will return 

3 < 2

```


################################
# Variable Assignment
################################

### A variable allows you to store a value (e.g. a number) or an object (e.g. a function description). 
### Later you can easily access the value or object stored within this variable
### Reminder: R is case sensitive

```{r}
# Example: Assign the value 25 to x

x <- 25

# Print out the value of the variable x
x
```

### You can view your variables in your Environment tab in RStudio

```{r}
# (s8) Assign the value of the square root of 144 to the variable my_var

my_var = sqrt(144)

# print out my_var

my_var

```

```{r}
# (s9) Assign the value of the 10 to the variable y

y = 10

# print out y

y

```

###  You can use arthmetic operators and functions on variables if they are NUMBERIC variables

```{r}
# (s10) Subtract my_var from y and divide by 4 (Hint: your answer should be -0.5)

(y - my_var)/4


```

```{r}
# (s11) Assign the above equation to the variable my_ans and print it out

my_ans = (y-my_var)/4

```

```{r}
# (s12) Assign the string "hello world" to the the variable char
char= "hello world"

```

```{r}
# (s13) Add char and y together (Hint:you will get an error, why?)

#y + char

# trying to add two different types

```

#############################
# Basic Data Types
############################

* Decimal values like 5.88 are called numerics
* Whole numbers like 5 are called integers. Integers are also numerics
* Boolean values (e.g., TRUE or FALSE) are called logical
* Text (or string) values are called characters (e.g. your variable 'char')

### We can check the data type of variable using the class() function

```{r}
# Example: Declare variables of different types

numeric <- 25
character <- "data science"
logical <- TRUE #note the color change of the text

```

```{r}
# (s14) Check the class of variables above using class() (Hint: class(numeric))

class(numeric)
class(character)
class(logical)

```

#############################
# Vectors
#############################

### Vectors are one-dimension arrays that can hold numeric, character, or logical data
### In R, you use the combine function, c(), with the elements separated with commas

```{r}
# Example: Creating vectors
numeric_vector <- c(41, 66, 23, 3)

# print numeric vector
numeric_vector

# check the class
class(numeric_vector)
```

### Vector Selection
* use square brackets ([]) to select a specific element(s) out of your vector
* note: unlike other programming languages, the first element in a R vector is 1 (compared to 0)

```{r}
# Example
streets <- c('ivy', 'emmet', 'jpa', 'main')
streets[2]
```

```{r}
# (s18) Use your days_name vector and select the element "Thursday"

days_name = c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday')


days_name[4]

```

```{r}
# Example: Select multiple elements

# select the first 3 elements
streets[c(1, 2, 3)]
# =
streets[1:3]

# select the first and last elements (note the use of c())
streets[c(1,4)]

# select the first 2 elements and the last element
streets[c(1:2, 4)]
```

### Element selection can also be done using the names of elements instead of their numeric positions

```{r}
#Example:
# Vector of street numbers
street_numbers <- c('Street 1', 'Street 2', 'Street 3', 'Street 4')

# Assign street numbers as names to Streets
names(streets) <- street_numbers

# Select Street 1
streets['Street 1']

# Select Street 1 and Street 4
streets[c('Street 1', 'Street 2')]
```


```{r}
# (s15) Create, print, and check the class of a vector of your three favorite TV shows (Hint: don't forget the quotation marks) and assign it the variable tv

tv = c('Game of Thrones', 'Attack on Titan', 'Death Note')

class(tv)

tv

```
