---
title: "Assignment - Shiny"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float: yes
    code_folding: show
#runtime: shiny    
---

- You may work up to 3 classmates on this assignment. For Part 2, please use a different dataset from those you are working with on the term project. 
- Upload your Shiny app on [ShinyApps.io](https://www.shinyapps.io).

- Submit your `.Rmd` AND your URL(s) (https://XXXXX.shinyapps.io/XXXXX) on Canvas.

- I recommend building your apps locally and deploying them to Shiny after the build to conserve your free hours on shinyapps.io

```{r}
# List classmates you worked with on this homework:

# 1. wfl9zy

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(knitr)
library(tidyverse)
library(plotly)
library(foreign)

```

# Part 1 (10 points)

- First, get the latest reported counts of COVID-19 cases and deaths from WHO:[Data](https://covid19.who.int/data).
      
- The file should be named `WHO-COVID-19-global-table-data.csv`
  
- Create a subset including 3 countries per `WHO region` (Africa, Americas, Eastern Mediterranean, Europe, South-East Asia, Western Pacific). You can choose any three countries within each `WHO region` to compare the mortality rate (`mutate(rate = "Deaths - cumulative total"/"Cases - cumulative total")`). 

- Create a shiny app to produce bar plots, along with the control type of `selectInput()`.

>You need to use `selectInput()` in the `ui` to select one level from the `WHO region` variable (with 6 levels: Africa, Americas, Eastern Mediterranean, Europe, South-East Asia, Western Pacific). 
     
- Your output should be a barplot of mortality rates from three countries in the selected region (Error bars not required)


```{r}
# add your codes 

library(readr)

data <- read_csv("./data/WHO-COVID-19-global-data.csv")

  
library(shiny)
library(dplyr)


ui <- fluidPage(
  selectInput("Region", "Select a WHO Region", choices=unique(data$WHO_region)),
  plotOutput("barplot")
)

server <- function(input, output) {
  output$barplot <- renderPlot({
    selectedRegion <- input$Region
    selectedData <- data %>%
      filter(WHO_region == selectedRegion) %>%
      group_by(Country) %>%
      summarize(rate = mean((New_deaths - Cumulative_deaths) / (New_cases - Cumulative_cases), na.rm = TRUE))
    
    
    preparedData <- selectedData %>%
      top_n(3, rate)
    
    ggplot(preparedData, aes(x=Country, y=rate, fill=Country)) + 
      geom_bar(stat="identity") + 
      labs(
        x= paste("Top 3 Covid Death Rate Countries in ",selectedRegion),
        y="Covid Death Rate",
        title=paste("Top 3 Covid Death Rates in ",selectedRegion)
      )
    
    
  })
}


shinyApp(ui, server)

```

# Part 2 (10 points)

- Use a dataset that is available in data repositories (e.g., [kaggle](https://www.kaggle.com/datasets)). 
- e.g: "Top 50 US Tech Companies 2022 - 2023.csv"

- Select *two* continuous variables (denoted as "A" and "B") 

- Create an interactive scatterplot between "A" and "B" with a loess line using Shiny app. 

- You need to create two input variables for loess regression. 

    - One input variable is the degree with values of 0, 1, and 2.
    - The other input variable is the span value ranging from 0.2 to 0.9.
   
   
- You can choose appropriate control types from [Widget gallery](https://shiny.rstudio.com/gallery/widget-gallery.html).
     
```{r}

data_mcdonalds <- read.csv("./data/mcdonalds_dataset.csv")

data_mcdonalds_clean <- data_mcdonalds %>%
  mutate(product_calories = as.numeric(sub(".*: ", "", product_calories)))

data_mcdonalds_clean <- data_mcdonalds_clean %>%
  mutate(product_price = as.numeric(sub("£ ?", "", product_price)))


data_mcdonalds_final <- na.omit(data_mcdonalds_clean)


ui <- fluidPage(
  titlePanel("Interactive Scatterplot with Loess Line"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("span", "Span Value:", min = 0.2, max = 0.9, value = 0.5),
      selectInput("degree", "Degree:", choices = c(0, 1, 2), selected = 1)
    ),
    mainPanel(
      plotOutput("scatterplot")
    )
  )
)

server <- function(input, output) {
  output$scatterplot <- renderPlot({
    degree <- input$degree
    span <- input$span

    # Create the scatterplot with a loess line
    ggplot(data_mcdonalds_final, aes(x = product_price, y = product_calories)) +
      geom_point() +
      geom_smooth(method = "loess", formula = y ~ x, degree = degree, span = span) +
      labs(
        title = "Scatterplot with Loess Line Of Product Price by Product Calories at Mcdonalds",
        x = "Product Price",
        y = "Product Calories"
      )
  })
}

shinyApp(ui, server)

```

- Describe a relationship between A and B. 

## We can clearly see a positive correlation between price and calories. As you pay more, you get more food!
