---
title: "Report Clone"
author: "wlf9zy"
date: "2023-10-04"
output: 
  html_document:

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Necessary Packages: 

```{r}
library(readr)
library(tidyverse)
```


# Duplicated Report! My Data Below:

# Load Data:

```{r}
flights <- read.csv("../data/flights.csv")
airports <- read.csv("../data/airports.csv")
```

# Datasets:

## **Flights**

**A dataset with information about individual flights including: `DayofMonth`, `DayofWeek`, `Carrier`, `OriginAirportID`, `DestAirportID`, `DepDelay`, `ArrDelay`**

```{r}
head(flights)
```

## **Airports**

A dataset with the `airport_id`, `city`, `state`, and `name` for 365 airports


```{r}
head(airports)
```

# Report Objectives:

1. Determine how many airports are in New York, New Jersey, and Connecticut (the tri-state area).
2. Visualize the average arrival and departure delays at each airport identified in Objective 1.
3. Analyze which carriers have the most and least departure and arrival delays.

# Objective 1:

**Determine how many airports are in New York, New Jersey, and Connecticut (the tri-state area).**

```{r}
tristate_airports <- filter(airports, state=="NY",state=="NJ", state=="CT")

n_distinct(tristate_airports$airport_id)

```

# Objective 2:

**Visualize the average arrival and departure delays at each airport identified in Objective 1**

```{r}
flights <- flights %>%
  pivot_longer(cols = c('DepDelay', 'ArrDelay'), names_to = 'delay_type', values_to = 'delay_min')

head(flights)
```

```{r}
colors = c(ArrDelay = '#a8dadc', DepDelay = '#219ebc')

p <- ggplot(flights, aes(delay_type, delay_min, fill = delay_type)) +
  geom_bar(position = 'dodge2', stat = 'summary', fun = 'mean') +
  scale_fill_manual(values = colors) +
  ylab('Minutes') +
  xlab('Delay Type') +
  labs(
    title = "Average Delay by Departure and Arrival"
  ) + theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position="none") +
  coord_flip()

p
```

<center>thank you</center>

