---
title: 'DS2003: Interactive Plots in plotly'
author: "Natalie Kupperman"
week: 9
class: 17
---
## Plotly

* {plotly} is an open source graphing library 

* `plotly` is a high-level interface to [plotly.js](https://github.com/plotly/plotly.js), based on [d3.js](https://d3js.org/) which provides an easy-to-use UI to generate slick D3 interactive graphics.

* Plot work in multiple formats: viewer windows, R Markdown documents, shiny apps

* plotly is also backed by a strong community and active development

* plotly works with R, python, excel, and others. We will just focus on R in this class.  

### Why use interactive graphics?
  - Zooming
  - Hovering
  - Sliding
  - Filtering
  
### Difference between interactive and animated graphs
- interactive graphs allow the user to manipulate different features of the graph

  
### plotly Resources

- [The Plolty Website](https://plotly.com/)
- [The Plotly R documentation](https://plotly.com/r/)
- [The Plotly R Package on GitHub](https://github.com/plotly/plotly.R)
- [The Plotly R Cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf)

```{r}
library(plyr) # https://cran.r-project.org/web/packages/plyr/index.html
library(tidyverse)
library(plotly) # (https://plotly.com/)
```

```{r}
# read in 'wine.csv' data
wine <- read.csv('wine.csv')

head(wine)
```
```{r}
wine$Type <- as.factor(wine$Type)

```


## Convert Static Graphs to Interactive Graphs Using `ggplotly`

Let's look at a static graph comparing `flavanoids` to `proline`

```{r}
static_plot <- wine %>%
  ggplot(aes(x = Flavanoids, y = Proline, color = Type)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))

static_plot
```

### How to create a plotly object

- There are two main approaches to initialize a plotly object:  
  
   + transforming a ggplot2 object with `ggplotly()` 
   + setting up aesthetics mappings with `plot_ly()` directly
     * Use ggplot style attribute to subset the data (can be harder to control attributes)
     * Manually subset data for each trace (maximum control of attributes)
     

## Today's focus `ggplotly()`

- `ggplotly()` takes existing `ggplot2` objects and converts them into interactive `plotly` graphics. That is, `ggplotly()` converts your static plots to an interactive web-based version!

- This makes it easy to create interactive figures because we are already familiar to the `ggplot2` syntax.

```{r}
# Create an interactive plot of Flavonoids vs Proline

ggplotly(static_plot)
```

### Notes

### toolbar from left to right are as follows:
Downloading the plot as a png file.
Zoom
Panning across the map. 
Selecting all points using a box. 
Selecting all points using a lasso. 
Zooming in and out on our plots. 
Resetting the view.
data

Interactive plots are great!  
*Bad Design = Bad Interactive Plots  
*Follow data-viz best practices  

### Bar Plot

```{r}
g <- ggplot(data = wine, aes(x = Type, fill = Type)) + 
     geom_bar(fill = c("#00AFBB", "#E7B800", "#FC4E07")) +
     theme_classic() 

ggplotly(g)
```

### If you want the bar in descending order
```{r}
g.5 <- count(wine, Type) %>% 
       ggplot(aes(x = reorder(Type, -n), y = n)) +
       geom_bar(fill = c("#00AFBB", "#E7B800", "#FC4E07"), stat = 'identity') +
       theme_classic() +
       xlab('Type')
     
ggplotly(g.5)
```

### Histogram

```{r}
g2 <- ggplot(wine, aes(x = Flavanoids, fill = as.factor(Type))) + 
      geom_histogram(bins = 15, alpha = 0.6) +
      theme_minimal() 
      scale_fill_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))

ggplotly(g2)

```

### Scatterplot

```{r}
g4 <- ggplot(wine, aes(x = Flavanoids, y = Proline, color = as.factor(Type))) + 
      geom_point() + 
      geom_smooth(method="loess", formula=y~x, se=F) + 
      scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"), name = 'Type') +
      theme_bw() 

ggplotly(g4)

# Flavonoids are rich in antioxidant
# A large percentage of the total amino acid left in the wine is in the form of proline

```




## Creating interactive graphs directly in `plot_ly`

- `plot_ly()` is the base `plotly` command to initialize a plot from a dataframe, similar to `ggplot()` from ggplot2.

### import `plotly`

```{r}
library(plotly)
```

### Bar plot
```{r}
wine %>% 
  count(Type) %>% #using dplyr create frequency table
  plot_ly(x = ~Type, y = ~n) %>% # initialize graph in plotly, use ~ for aes mappings
  add_bars() %>% #set graph as bar chart
  layout(title = "Freq. of Wine Soil Type",
         xaxis = list(title = "Type"),
         yaxis = list(title = "Frequency"))
```

### Reorder columns

```{r}
wine %>% 
  count(Type) %>% #using dplyr create frequency table
  mutate(Type = fct_reorder(Type, n, .desc = TRUE)) %>% #order for largest to smallest
  plot_ly(x = ~Type, y = ~n) %>% # initialize graph in plotly, use ~ for aes mappings
  add_bars() %>% #set graph as bar chart
  layout(title = "Freq. of Wine Soil Type",
         xaxis = list(title = "Type"),
         yaxis = list(title = "Frequency"))
```

### Basic Histogram
```{r}
g6 <- wine %>% plot_ly(x = ~ Flavanoids, type = "histogram")

g6
```

### Flavanoids by Type Histogram

## Explanation!

- This table creates and overlays three different types of wine, these represent the three different colors. 
- The y-axis is the count of wines of each type of soil. 
- This graph fits information about three different bins, one for each soil type, all into one convenient super-imposed visual.

```{r}
# Lets split up the data to illustrate the manual way of making the plots
Type1 <- wine %>% filter(Type == 1)
Type2 <- wine %>% filter(Type == 2)
Type3 <- wine %>% filter(Type == 3)

# Full Manual method without for loop, subsetting data before
g7a <- plot_ly(alpha = 0.4) %>%
  add_histogram(x = ~Type1$Flavanoids, name = 'Type 1', opacity = 0.6, 
                marker = list(color = "#00AFBB",
                              alpha = 0.6,
                              line = list(color = "lightgray",
                              width = 2))) %>%
  add_histogram(x = ~Type2$Flavanoids, name = 'Type 2', opacity = 0.6,
                marker = list(color = "#E7B800",
                              line = list(color = "lightgray",
                              width = 2))) %>%
  add_histogram(x = ~Type3$Flavanoids, name = 'Type 3', opacity = 0.6,
                marker = list(color = "#FC4E07",
                              line = list(color = "lightgray",
                              width = 2))) %>%
  layout(barmode = 'overlay',
         title = 'Histogram of flavanoids by soil type',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'count'))

g7a

```


## Explanation!

- This graph uses a for loop to iterate over adding each histogram, dont repeat yourself!

```{r}
# Manual method with for-loop that temp subsets data by soil type on the fly
my_colors <- c("#00AFBB","#E7B800","#FC4E07")
g7b <- wine %>% 
  plot_ly() # initialize graph in plotly, use ~ for aes mappings

  for (n in 1:(length(levels(wine$Type)))) {
    temp <- subset(wine, Type == levels(wine$Type)[n])

    g7b <-  add_histogram(g7b, data = temp, x = ~Flavanoids, name = paste0("Type ", n), 
                          opacity = 0.6, marker = list(color = my_colors[n], line = list(color = "lightgray",
                              width = 2)))
  }
 g7b <- g7b %>% layout(barmode = 'overlay',
         title = 'Histogram of flavanoids by soil type',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'count'))
g7b

```



## Explanation!

- This graph uses built in plotly functionality to automate the entire process! 


```{r}
# Automated method, ggplot style
g7c <- wine %>% 
  # initialize graph in plotly, use ~ for aes mappings
  plot_ly(x = ~Flavanoids, color = ~Type, colors = c("#00AFBB","#E7B800","#FC4E07")) %>% 
  
 add_histogram(opacity = 0.6, marker = list(line = list(color = "lightgray", width = 2))) %>%
  
 layout(barmode = 'overlay',
         title = 'Histogram of flavanoids by soil type',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'count'))
g7c


```


## Scatter plot

## Explanation!

- This visual graphs a scatter plot of flavinoids and proline and then sets the color of the point to be based on the soil type! 

```{r}
# Manual subset method with more control for formatting
# Using dataframe subsets Type1, Type2, Type3 we generated earlier
g8a <- plot_ly(type = 'scatter', mode = 'markers')  %>%
  add_trace(x = Type1$Flavanoids,y = Type1$Proline,opacity = 0.5, 
      marker = list(color = "#00AFBB", size = 7,line = list(color = 'lightgray',
        width = .5)),name = 'Type 1') %>%
  
  add_trace(x = Type2$Flavanoids,y = Type2$Proline,
    marker = list( color = "#E7B800", size = 7,line = list(color = 'lightgray',
        width = .5)), name = 'Type 2')  %>%
  
  add_trace(x = Type3$Flavanoids, y = Type3$Proline, 
      marker = list(color = "#FC4E07",size = 7, line = list(color = 'lightgray',
        width = .5)), name = 'Type 3') %>%
  
  layout(title = 'Scatterplot of Flavanoids & Proline',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'proline'))
g8a
```


## Explanation!

- This plot once again leverages built in plotly functionality to accomplish the same task more easily.


```{r}
# Automated method, ggplot style
g8b <- wine %>% 
  plot_ly(x = ~Flavanoids, y = ~Proline, color = ~Type, colors = c("#00AFBB", "#E7B800", "#FC4E07"),
          type = 'scatter', mode = 'markers') %>%  # initialize graph in plotly, use ~ for aes mappings
 layout(title = 'Scatterplot of Flavanoids & Proline',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'proline'))

g8b
```




### 3D Scatterplot


## Explanation!

- This plot graphs three different axis to create a 3d manipulatable visual!

```{r}
g9a <- plot_ly(type = 'scatter3d', mode = 'markers') 
g9a <- g9a %>%
  add_trace(x = Type1$Flavanoids, y = Type1$Proline, z = Type1$Alcalinity, opacity = 0.5, 
    marker = list( color = "#00AFBB", size = 7,
      line = list(color = 'lightgray', width = .5)), name = 'Type 1')  %>%
  
  add_trace( x = Type2$Flavanoids, y = Type2$Proline, z = Type2$Alcalinity,
    marker = list( color = "#E7B800", size = 7, 
                   line = list( color = 'lightgray',width = .5)),name = 'Type 2') %>%
  
  add_trace( x = Type3$Flavanoids, y = Type3$Proline, z = Type3$Alcalinity,
    marker = list( color = "#FC4E07", size = 7,
      line = list( color = 'lightgray', width = .5 )), name = 'Type 3') %>%
  
  layout(title = 'Scatterplot of Flavanoids, Proline & Alcalinity',
         scene = list(xaxis = list(title = "flavanoids"),
                      yaxis = list(title = "proline"),
                      zaxis = list(title = "alcalinity")))
g9a
```


## Explanation!

- This plot accomplishes the same task, but uses different colors and a more streamlined plotly approach to visual generation

```{r}
g9b <- wine %>% plot_ly( opacity = 0.5, type = 'scatter3d', mode = 'markers') %>% 
  add_trace(x = ~Flavanoids, y = ~Proline, z = ~Alcalinity, color = ~Type,
            marker = list(size = 7,line = list(width = .5)))  %>%
  
  layout(title = 'Scatterplot of Flavanoids, Proline & Alcalinity',
         scene = list(xaxis = list(title = "flavanoids"),
                      yaxis = list(title = "proline"),
                      zaxis = list(title = "alcalinity")))
g9b
```



## Line Plot
(From plotly examples)

## Explanation!

- This plot creates a scatter plot but uses the lines mode to draw a line through the data for each type in the ~supp column of the data

```{r}

# Obtain data from subset of tooth growth dataset
tg <- ddply(ToothGrowth, c("supp", "dose"), summarise, length=mean(len))


# Create plotly figure
g10 <- plot_ly(tg, x = ~dose, y = ~length, type = 'scatter', mode = 'lines', 
               linetype = ~supp, color = I('black'))  %>% 
  layout(title = 'The Effect of Vitamin C on Tooth Growth in Guinea Pigs by Supplement Type',
         xaxis = list(title = 'Dose in milligrams/day'),
         yaxis = list (title = 'Tooth length'))

g10
```

