---
title: "Quiz4"
author: "wlf9zy"
date: "2023-10-13"
output: html_document
---


# Quiz 4 (wfl9zy)

# Load Data and Import Libraries:
```{r}
library(tidyverse)
library(nycflights13)
library(ggplot2)
library(ggthemes)

# load the data:
data("flights")
```

# Part 1:

```{r}
flights_dat <- flights %>%
  inner_join(airlines, by = "carrier") %>%
  rename(airline_name = name)
```


## Clean Data For Q1:
```{r}
flights_summary <- flights_dat%>%
  group_by(airline_name, origin)%>%
  summarize(num_flights = n())
head(flights_summary)
```
# Question 1:
```{r}
ggplot(flights_summary, aes(x=airline_name, y=num_flights, fill=origin)) +
  geom_bar(stat="identity", position="dodge") +
  labs(
    x="Airline", 
    y="Number of Flights", 
    fill="origin",
    title="The Distribution of Flights by Airline and NYC Airport") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position="top")
```


# Question 2:


## Answer: The above change centered the title text directly above the visual!
```{r}
ggplot(flights_summary, aes(x=airline_name, y=num_flights, fill=origin)) +
  geom_bar(stat="identity", position="dodge") +
  labs(
    x="Airline", 
    y="Number of Flights", 
    fill="origin",
    title="The Distribution of Flights by Airline and NYC Airport") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position="top") + 
  theme(plot.title = element_text(hjust=0.5))
```


# Question 3:

## NOTE: In the visual displayed for the quiz, the colors do not match the hints, I followed the hints with JFK=tomato, LGA=turquoise, EWR=skyblue

## Generate Plot:

```{r}
ggplot(flights_dat, aes(x = origin, y = air_time, fill = origin)) +
  geom_boxplot(outlier.color = "purple", na.rm = TRUE) +
  scale_fill_manual(
    values = c("JFK" = "tomato", "LGA" = "turquoise", "EWR" = "skyblue")) +
  labs(
    x = "Origin",
    y = "Air Time",
    title = "Distribution and Quantiles of Flights by Origin"
  ) +
  scale_x_discrete(limits = c("JFK", "LGA", "EWR"))  # Order of boxplots
```

# Question 4:

## Answer: The theme_classic removed the background grey grid to give the visual a cleaner and more sleak look.

```{r}
ggplot(flights_dat, aes(x = origin, y = air_time, fill = origin)) +
  geom_boxplot(outlier.color = "purple", na.rm = TRUE) +
  scale_fill_manual(
    values = c("JFK" = "tomato", "LGA" = "turquoise", "EWR" = "skyblue")) +
  labs(
    x = "Origin",
    y = "Air Time",
    title = "Distribution and Quantiles of Flights by Origin"
  ) +
  scale_x_discrete(limits = c("JFK", "LGA", "EWR")) +  # Order of boxplots
  theme_classic()
```


# Trying some of the other themes!

```{r}
ggplot(flights_dat, aes(x = origin, y = air_time, fill = origin)) +
  geom_boxplot(outlier.color = "purple", na.rm = TRUE) +
  scale_fill_manual(
    values = c("JFK" = "tomato", "LGA" = "turquoise", "EWR" = "skyblue")) +
  labs(
    x = "Origin",
    y = "Air Time",
    title = "Distribution and Quantiles of Flights by Origin"
  ) +
  scale_x_discrete(limits = c("JFK", "LGA", "EWR")) +  # Order of boxplots
  theme_economist()

```





