---
title: "Quiz2"
author: "wfl9zy and gwu8ek"
date: "2023-09-15"
output: html_document
---


# Group wfl9zy and gwu8ek

```{r}

library(rvest)
library(dplyr)

premier_league <- read_html("https://www.premierleague.com/tables")

```


# Getting Club Names For Each League
```{r}
club_names_PL <- premier_league%>%
  html_nodes(".long")%>%
  html_text()

```

# Games Won At Each League

```{r}

games_won_PL <- premier_league%>%
  html_nodes("td:nth-child(4)")%>%
  html_text()

```

# Games Drawn At Each League

```{r}

games_drawn_PL <- premier_league%>%
  html_nodes("td:nth-child(5)")%>%
  html_text()
```

# GF Per League

```{r}

GF_PL <- premier_league%>%
  html_nodes(".hideSmall:nth-child(7)")%>%
  html_text()

GF_PL <- GF_PL[GF_PL!="GF"]

```


# Combine Data Into 3 DataFrames, one for each League

```{r}

epl_table <- data.frame(clubNamePL=club_names_PL, winCountPL=games_won_PL, drawCountPL=games_drawn_PL,GFCountPL=GF_PL)

head(epl_table)
```

# Converting to Numeric 

```{r}

epl_table$winCountPL <- as.numeric(epl_table$winCountPL)
epl_table$drawCountPL <- as.numeric(epl_table$drawCountPL)
epl_table$GFCountPL <- as.numeric(epl_table$GFCountPL)


glimpse(epl_table)

```

# Adding the League Column:

```{r}

epl_table2 <- epl_table


epl_table2$league <- 3 * epl_table$winCountPL + 1 * epl_table$drawCountPL



head(epl_table2)

```
# Question 3:

```{r}


epl_table3<-epl_table2%>%
mutate(league=c(rep("Premier League",20),
rep("First Division",26),
rep("U18",26)))%>%
slice(1:46)

head(epl_table3)


```

# Description:
- This function adds a League Column and populates the first 20 with Premier League and the back 26 with "First Division". It then Slices the Dataframe grabbing rows 1->46.


# Question 4:
```{r}

epl_premier <- epl_table3[epl_table3$league=="Premier League",]

median_premier <- median(epl_premier$GFCountPL)

epl_first_division <- epl_table3[epl_table3$league=="First Division",]

median_first_division <- median(epl_first_division$GFCountPL)


print(median_premier)
print(median_first_division)

```



